---
title: "Fuzzy Matching"
author: "Pedram Roshdinavid"
date: "August 30, 2016"
output: html_document
---

```{r setup, include=FALSE}
ROOT_DIR = rprojroot::find_root("README.md")
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = ROOT_DIR)
```

From David Doherty: https://connect.fg.rbc.com/message/587343#587343

I have a database of athletes for triathlons with all their results stored in 
them covering the last few years I will get lists of athletes for future events
and I want to cross compare them,
however some of the fields are not normalised

E.g. the names may not be expressed the same way 'John Smith' may be 'Smith, John'
or 'John J. Smith'. I want to employ fuzzy matching to cross compare them. 
In Excel this is horrific as I have to find and install an add-in and then once 
I email the sheet to someone, either the sheet doesn't work or they have to also 
install the add-in

There are additional helper columns to avoid mistmatches: e.g. you can identify
their age group; and; in previous results you are given their home town and country,
and in the future list you are given their state.

Once I've managed to cross-compare the results I want to run stuff like:
For everyone aged 25-29 in an upcoming event, show me their results where available, 
and order them by finishing time From the data below you should be able to see that 
'Tom Aaker' is doing an upcoming race, and this is the same guy that is listed as 'Thomas Aaker' in a previous result.

Eventually after that you may want to run additional more advanced analytics:
Normalise results for difficulty (e.g. by using the winners time as a scaling
mechanism for previous results)
Run simulations to figure out if athletes perform comparitively better on 
certain types of courses
 
Example data, results from previous races:
 
   
Name	State	Country	Gender	Division	Overall
Alicia Kaye	Clermont	USA	Female	FPRO	4:27:39
Ryan Ung	Los Angeles	USA	Male	M25-29	4:30:40
Lauren Barnett	Plano	USA	Female	FPRO	4:31:00
Aaron Church	South Riding	USA	Male	M40-44	4:31:43
Anthony Chan	Flushing	USA	Male	M18-24	4:35:03
Dan Stephens	Boise	USA	Male	M50-54	4:35:16
Thomas Aaker	San Francisco	USA	Male	M25-29	4:36:57
Jacob Slife	West Point	USA	Male	M18-24	4:37:43
Matt Smith	Denver	USA	Male	M35-39	4:38:07
Bill Picciano	Jupiter	USA	Male	M45-49	4:39:50
Steve Chaid	Wien	USA	Male	M45-49	4:39:58
   
 
Example for name for upcoming races
 
   
Name	Age group	State
Tom Aaker	M25-29	CA
James Aaron	M35-39	AL
Dave Abel	M45-49	DC
Robert Abernathy	M45-49	GA
Richard Abrahamson	M35-39	MN
Sidney Abramson	M25-29	KY
Filip Achten	M30-34	NY
Luis Acuna	M30-34	IL
Elizabeth Acuna	F30-34	IL
Jennifer Adamo	F40-44	IN

# Fuzzy Matching

This might be the more difficult part of this job, but not impossible. 
First, let's import some data from the web

```{r, message = F}
library(readxl)
library(dplyr)
tri <- read_excel("practical_examples/itu-world-triathlon-series-male.xls")
head(tri)
```

Now, let's create a full name field 

```{r}
tri <- tri %>%
  mutate(full_name = paste(`First Name`, `Last Name`))

head(tri[,c(4,5,17)])

```

Now let's create a fake name and see if we can match it

```{r}
fns <- c('Mario Molla', "Molla, Mario", "Mola, Mario", "Mario Mola Jr.", "Catherine Great", "John Travolta")
rn <- 'Mario Mola'

library(stringdist)
stringdist(fns, rn) # Default method
stringdist(fns, rn, method = 'qgram') # Seems to work better for our examples

```

Now that we have a method, let's create a name and find the best match in our 
dataset
```{r}

m <- stringdist('Pedram Navid', tri$full_name, method = 'qgram')
m
min(m)
# Best match is a 8. Let's find it
which.min(m) 

# 96th one! 
tri$full_name[96]

# Not very close. Let's try another name
m <- stringdist('Greg Billington', tri$full_name, 'qgram')
min(m)
which.min(m) 
tri$full_name[49]

# Pretty good! So, what threshold do we use?
methods <- c('osa', 'lv', 'dl', 'hamming' , 'lcs', 'qgram', 'cosine', 'jaccard', 'jw')
for(i in methods) {
  m <- stringdist('Billington, Greg', tri$full_name, method =  i)
  print(paste("Method:", i, tri$full_name[which.min(m)]))
}

# make a function to find the best method for a given name

find_best <- function(name) {
  methods <- c('osa', 'lv', 'dl', 'hamming' , 'lcs', 'qgram', 'cosine', 'jaccard', 'jw')
  for(i in methods) {
    m <- stringdist(name, tri$full_name, method =  i)
    cat(paste0(i, "\t  ", tri$full_name[which.min(m)], "\t", round(min(m), 2)))
    cat('\n')
  }
}
find_best("Sissons Jr., Ryan")
find_best("Donald Alfredon")
find_best("Aron Roile")
find_best("Matt Sharpe")

```


# Analysis of results 
For everyone aged 25-29 in an upcoming event, show me their results where available, and order them by finishing time

```{r}
library(ggplot2)
# create approx. age column and get rid of unneeded columns
tri_small <- tri %>%
  mutate(age = 2016 - YOB) %>%
  select(rank = Ranking, full_name, country = Country, age, events = `Total Events`,
    wcups = `World Cups`, points = Points, best_score = `Current Best Score`)
head(tri_small)

tri_small %>%
  filter(age>=25, age<=29) %>%
  top_n(5, best_score)

tri_small %>%
  filter(age>=35) %>%
  top_n(5, points)

ggplot(tri_small, aes(age, best_score, color = events)) +
  geom_point() + 
  geom_smooth()
```