# Subsetting


From David Doherty: https://connect.fg.rbc.com/message/587342#587342

Another work example:
 
You have a massive table of data in Excel, and you want to do processing on a
subset of that info. Normally I would autofilter the table and then select the 
data I'm interested in, the view changes to only show that information. 


# Get the data ready

We'll create some fake data based on the example provided. Here we pick a random
sample from a list of currencies and aliases and create a dataset with 100 
observations

```{r, message = FALSE, warning = FALSE}
library(dplyr)
library(ggplot2)
library(stringr)
library(knitr)

# Set the seed so we get the same result every time
set.seed(1234)

# Let's create the dataset from the example first
trade_currencies <- c('EUR', 'JPY', 'CAD', 'USD')
trade_alias <- c('InflationAmortCapFloorType2', 'ZCInflationSwap', 'FXOption',
  'IR_BERMUDAN_SWAP', 'CMS_CAPFLOOR', 'CMS_SPREAD_SWAP', 'SomethingElse',
  'ThisOtherThing')

trade_id <- seq(1000, length.out = 100)
col1 <- sample(trade_currencies, size = 100, replace = TRUE)
col2 <- sample(trade_alias, size = 100, replace = TRUE)

dataset <- data.frame(id = trade_id, currency = col1, alias = col2)

head(dataset)
```


# Manipulate the data 

Let's tag based on the alias. First we will convert everything to uppercase
to make it easier to search for variation in capitalization 

```{r}

dataset_tag <- dataset %>%
  mutate(alias = str_to_upper(alias),  # Convert to upper
    cap = str_detect(alias, 'CAPFLOOR'), # add tags to new variables
    swap = str_detect(alias, 'SWAP'),
    inf = str_detect(alias, 'INFLATION'),
    spr = str_detect(alias, 'SPREAD'),
    tag_count = cap + swap + inf + spr # Count the number of 'hits'
  )

head(dataset_tag)
```


# Easy filters
```{r}
filter(dataset_tag, cap == TRUE, currency == "USD")

filter(dataset_tag, cap, currency == "USD") # Equivalent 

filter(dataset_tag, !swap, currency == "JPY") %>%
  head()

filter(dataset_tag, swap, tag_count > 1) %>%
  head() %>%
  kable()


```