---
title: "Visualizing Data"
author: "Pedram Navid"
date: "August 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Visualizing data is where you'll find the most bang for your R-buck and so I've
decided to start there rather than with the more mundane, tedious, and frankly
boring tasks of importing and cleaning data.

That's not to say that those tasks aren't important, but hopefully by starting
with the fun stuff, I'll win you over, and once it's time to discuss importing
Excel files and doing data quality checks you'll stick around.

There's more than one way to get started with visualization: there's base R
graphics, lattice, ggplot2 and others. In line with this course's philosophy
we'll be using ggplot2, which is part of the *tidyverse* of packages (these
are packages that work well together and share a common philosophy) mostly
authored by Hadley Wickham. Other packages in the tidyverse include dplyr and 
tidyr, and we'll be using those frequently as well. 

## Visual Exploration of Data

**ggplot2** has two main functions: **qplot** and **ggplot**. **qplot** is great
for initial data exploration to quickly see trends, distributions, and outliers
without focusing too much on visual presentation. *ggplot* is used once you'd
like to refine the visual display of data for sharing with others. 

```{r}
# mtcars was extracted from the 1974 Motor Trend US magazine, and comprises
# fuel consumption and 10 aspects of automobile design and performance for 
# 32 automobiles (1973–74 models).
library(ggplot2)
head(mtcars)

# What's the relationship between mileage and horsepower?
qplot(data=mtcars, mpg, hp)

# Is there a difference in that relationship when looking at automatic (0) or manual (1)?
qplot(data=mtcars, mpg, hp, colour = am)
```

Notice that ggplot thinks that the variable **am** that describes whether
a car is automatic or manual is a continous variable, when it is actually
categorical? Let's fix that by converting **am** into a factor:

```{r}
qplot(data=mtcars, mpg, hp, color = factor(am))
```

There we go. Now it's clear that **am** is a categorical variable. 

```{r}
# Break out the plots further by number of gears.
qplot(data=mtcars, mpg, hp, color = factor(am), facets = ~ gear)
```