---
title: "Visualizing Data"
author: "Pedram Navid"
date: "August 14, 2016"
output:
  ioslides_presentation: 
    fig_height: 3
    fig_width: 5
    widescreen: yes
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

# Overview

## Why Viz First?

Philosophy: start with the stuff you'll want

> - so you'll stick around for the stuff you hate.

## Visual Exploration of Data

- Most important part of data exploration
- Ability to see trends and relationships is unmatched by any statistical summary
- One of the most important methods of communication 
- Analysis often hinges on proper visual exploration

## Plotting Packages in R

- Lots of packages: base, lattice, ggplot2, others..
- In keeping with the philosophy of this tutorial, we will focus on ggplot2
- Two main functions: `ggplot()` and `qplot()`

```{r}
# Install if you haven't
# install.packages('ggplot2')
library(ggplot2)
```

# qplot()

## qplot: fast plotting

- qplot gives you fast plotting, at the expensive of customizability
- great for throwing ideas out quickly and exploring new possibilities

```{r, message = F, warning = F}
library(MASS)

# Histogram is default for one variable
qplot(Claims, data = Insurance)
```

## 

```{r}
# Scatterplot is default for two variables
qplot(Holders, Claims, data = Insurance)
```

Basic format: `qplot(x, y, data = Your_Dataset)`

## qplot: expanded

```{r}
# Add 3rd variable to aesthethics such as colour, size, fill
qplot(Holders, Claims, colour = Age, data = Insurance)
```

## 

```{r}
# Don't ever do this, it's very hard to interpret
qplot(Group, Claims, size = Holders, colour = Age, data = Insurance)
```

## qplot: facets

```{r}
# Facet breaks up a plot by a group.
qplot(Claims, facets = ~ Age, data = Insurance)
```

facets use the `formula` syntax: `~ a` for a single group or `a ~ b` for 2 groups

##
```{r}
qplot(Holders, Claims, facets = Group ~ Age, data = Insurance)
```

# ggplot()

## ggplot primer

Syntax is confusing at first.

```{r}
ggplot(data=Cars93, aes(x = Horsepower, y = MPG.city)) +
  geom_point()
```

Basic format is:  
`ggplot(data=YourDataSet, aes(x = IndVariable, y = DepVariable)) +
geom_XXXX()`

# ggplot: aesthethics

## aesthethics

* aesthethics are something that the plot draws that varies with data 
* examples: 
    + colour of a point or line
    + size of a point, or line
    + fill of a bar, histogram
    + shape of a point
*  use aes() to define them, either for the whole plot: `ggplot(data = bla, aes(x, y))`
* or for individual layers, if different layers have different aesthethics
    + `geom_line(aes(date, value1, colour = group)) + `  
  `geom_line(aes(date, value2, colour = group))`

# ggplot: aesthethics -- examples

## Colour
```{r}
# Position (required) + discrete colour aesthethic 
ggplot(data=Cars93, aes(x = Horsepower, y = MPG.city, colour = Origin)) +
  geom_point()
```

# Continous colour
```{r}
ggplot(data=Cars93, aes(Horsepower, MPG.city, colour = Weight)) + 
  geom_point()
```

## Size
```{r}
ggplot(data=Cars93, aes(Horsepower, MPG.city, size = Weight)) + 
  geom_point()
```

## Fill (histograms)
```{r}
ggplot(data=Cars93, aes(Horsepower, fill = Origin)) +
  geom_histogram(binwidth = 10)
```

## Fill (bars)
```{r}
ggplot(data=Cars93, aes(Type, fill = DriveTrain)) + 
  geom_bar()
```

## Shape
```{r}
# Not as useful as colour
ggplot(data=Cars93, aes(Horsepower, Weight, shape = Type)) +
  geom_point()
```

## Size
```{r}
# Size aesthethic
ggplot(data=Cars93, aes(Horsepower, Price, size = EngineSize)) + 
  geom_point()
```

## Alpha
```{r}
# You can combine multiple aesthethics for multi-dimensional data
# Careful, as 4 dimensions becomes very hard for others to interpret
ggplot(data=Cars93, aes(Horsepower, MPG.city, alpha = Weight, colour = Type)) +
  geom_point()
```

# ggplot: geoms

## Overview of geoms

A geom is a thing that ggplot draws based on data. It will manipulate the data
in some way (sometimes) and then draw it on a plot. 

The ggplot2 cheatsheet is really helpful here: https://www.rstudio.com/wp-content/uploads/2015/12/ggplot2-cheatsheet-2.0.pdf

# ggplot geoms: 1 variable (continous)

## histogram

```{r, message=TRUE}
# If you don't provide bins, ggplot2 will (rightly) complain
ggplot(Cars93, aes(Horsepower)) + 
  geom_histogram()
```

## with bins
```{r, message=TRUE}
ggplot(Cars93, aes(Horsepower)) + 
  geom_histogram(binwidth=20)
```

## freqpoly
```{r}
ggplot(Cars93, aes(MPG.city)) +
  geom_freqpoly(binwidth=10)
```

# ggplot geoms: 1 variable (discrete)

## bar charts

```{r} 
ggplot(Cars93, aes(Type)) +
  geom_bar()
```

## bar chart -- weighted
```{r}
ggplot(housing, aes(Type, weight = Freq)) +
  geom_bar()
```

## bar chart - identity
```{r}
ggplot(Cars93[1:5,], aes(Make, Horsepower)) + 
  geom_bar(stat="identity")
```

# ggplot geoms: 2 variables

## point / scatterplot
```{r}
ggplot(Cars93, aes(EngineSize, Cylinders)) + 
  geom_point()
```

## jitter
```{r}
ggplot(Cars93, aes(EngineSize, Cylinders)) + 
  geom_jitter()
```

## boxplot

```{r}
ggplot(Cars93, aes(Type, Horsepower)) + 
  geom_boxplot()
```

## line
```{r}
ggplot(economics, aes(date, pop)) + 
  geom_line()
```

## multiple lines
```{r}
ggplot(economics) + 
  geom_line(aes(date, psavert, color = "Personal Savings Rate")) + 
  geom_line(aes(date, uempmed, color = "Unemployment Rate"))
```

## smoothing/regression lines
```{r}
ggplot(Cars93, aes(MPG.city, Horsepower)) + 
  geom_point() +
  geom_smooth()
```

## linear lines, without standard error
```{r}
ggplot(Cars93, aes(MPG.city, Horsepower)) + 
  geom_point() + 
  geom_smooth(method="lm", se=FALSE)
```

# Your Turn