---
title: "Visualizing Data"
author: "Pedram Navid"
date: "August 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Visualizing data is where you'll find the most bang for your R-buck and so I've
decided to start there rather than with the more mundane, tedious, and frankly
boring tasks of importing and cleaning data.

That's not to say that those tasks aren't important, but hopefully by starting
with the fun stuff, I'll win you over, and once it's time to discuss importing
Excel files and doing data quality checks you'll stick around.

There's more than one way to skin a cat with visualization (sorry #rcatladies):
there's base R graphics, lattice, ggplot2 and others. In line with this course's
philosophy we'll be using ggplot2, which is part of the *tidyverse* of packages
(these are packages that work well together and share a common philosophy) mostly
authored by Hadley Wickham. Other packages in the tidyverse include dplyr and 
tidyr, and we'll be using those frequently as well. 

## Visual Exploration of Data

### qplot

**ggplot2** has two main functions: **qplot** and **ggplot**. **qplot** is great
for initial data exploration to quickly see trends, distributions, and outliers
without focusing too much on visual presentation. *ggplot* is used once you'd
like to refine the visual display of data for sharing with others. 

```{r}
# mtcars was extracted from the 1974 Motor Trend US magazine, and comprises
# fuel consumption and 10 aspects of automobile design and performance for 
# 32 automobiles (1973–74 models).
library(ggplot2)
head(mtcars)

# What's the distribution of hp look like
qplot(data=mtcars, hp) 

# Shrink the bins down to 25-hp increments
qplot(data=mtcars, hp, binwidth=25)

# What's the relationship between mileage and horsepower?
qplot(data=mtcars, mpg, hp)

# Is there a difference in that relationship when looking at automatic (0) or manual (1)?
qplot(data=mtcars, mpg, hp, colour = am)
```

Notice that ggplot thinks that the variable **am** that describes whether
a car is automatic or manual is a continous variable, when it is actually
categorical? Let's fix that by converting **am** into a factor:

```{r}
qplot(data=mtcars, mpg, hp, color = factor(am))
```

There we go. Now it's clear that **am** is a categorical variable. 

```{r}
# Break out the plots further by number of gears.
qplot(data=mtcars, mpg, hp, color = factor(am), facets = ~ gear)
```

## ggplot

qplot is *okay* but it's not well suited for analysis that will be shared, and if
analysis isn't being shared, then it's not very useful. 

ggplot is the stronger, beefier cousin of qplot. Let's start with a simple chart
and beef it up step by step.

```{r}
# This data set gives the measurements in cms of the variables sepal length 
# and width and petal length and width, respectively, for 50 flowers from each 
# of 3 species of iris. The species are Iris setosa, versicolor, and virginica.

summary(iris)

# Our first attempt at plotting Sepal Length
ggplot(data = iris, aes(x = Sepal.Length)) + 
  geom_histogram()

# Now let's fix the bins of the histogram
ggplot(data = iris, aes(x = Sepal.Length)) + 
  geom_histogram(binwidth=0.5)

# Let's add some labels
ggplot(data = iris, aes(x = Sepal.Length)) + 
  geom_histogram(binwidth=0.5) + 
  labs(x = "Sepal Length (cm)", y = "Count of Observations", 
         title = "Distribution of Sepal Length")

# Let's break it out by species
ggplot(data = iris, aes(x = Sepal.Length)) + 
  geom_histogram(binwidth=0.5) + 
  labs(x = "Sepal Length (cm)", y = "Count of Observations", 
         title = "Distribution of Sepal Length") + 
  facet_wrap(~ Species)

# Maybe one column will be easier to visualize
ggplot(data = iris, aes(x = Sepal.Length)) + 
  geom_histogram(binwidth=0.5) + 
  labs(x = "Sepal Length (cm)", y = "Count of Observations", 
         title = "Distribution of Sepal Length") + 
  facet_wrap(~ Species, ncol = 1)

```

Normally, you wouldn't code like I did above, and I won't do that again, but 
I wanted to show the progression of a simple plot to more and more complex as 
you start to formula what it is you might want to look at. 

It's good practice to always plot the data. Plot the data in as many different 
ways as you can to find interesting outcomes. You won't necessarily show every
attempt (in fact, you'll probably show 1% of what you actually do), but it's
import to explore to better understand what the data might be telling you.

Let's continue but with two variables

```{r}
ggplot(data = iris, aes(x = Petal.Length, y = Petal.Width)) +
  geom_point()

# Odd, there's two distinct clusters. What's going on here?
ggplot(data = iris, aes(x = Petal.Length, y = Petal.Width)) +
  geom_point(aes(colour = Species))

# Well that was easy. 
```

